### 2023-12-27 misc/expo/additive.tex
Lisa

\proclaim Lemma~\advthm. There exists a real constant $c>0$ such that the following holds.
Suppose that there is $N\in \NN$ and $A\subseteq [N]$ of density $\delta = |A|/N$ such that
$A-A$ does not contain any positive square numbers. Then there exists an arithmetic progression
$P$ of length $\lfloor N^{1/3}\rfloor$ such that $|A\cap P| \ge (\delta + c\delta^3)|P|$.

Suppose that for
some $\delta_0>0$ such that for arbitrarily large $N\in \NN$, we can find $A\subseteq [N]$
of size $|A|\ge \delta_0 N$ such that $A-A$ does not contain a square. Then for
arbitrarily large $N'\in \NN$, there exists a subset
$A'\subseteq [N']$ such that $|A'|\ge (\delta_0 + c{\delta_0}^3)N'$ and $A'-A'$ does not contain a square.

\proof Suppose $\delta_0>0$ is such that for arbitrarily large $N$ we can find $A\subseteq [N]$
of density $\delta = |A|/N \ge \delta_0$ such that $A-A$ does not contain a square. Then for all
$n\in [N]$, we have

@@

$$\ex_{n\in [N]} \ex_{f\in [N^{1/3}]} f(n+dr) = O(N^{-2/3+ 1/100}).$$

@@

>>>>

\proclaim Lemma~\advthm.
Suppose that there is $N\in \NN$ and $A\subseteq [N]$ of density $\delta = |A|/N$ such that
$A-A$ does not contain any positive square numbers. Then there exists an arithmetic progression
$P$ of length $\lfloor N^{1/3}\rfloor$ and step size at most $2N^{1/100}$
such that $|A\cap P| \ge (\delta + c\delta^3)|P|$, where $c$ does not depend on $\delta$ or $N$.

\proof By hypothesis, $n\in [N]$, we have

@@

$$\ex_{n\in [N]} \ex_{r\in [N^{1/3}]} f(n+dr) = O(N^{-2/3+ 1/100}).$$

@@

Recall that at the analogous point in the proof of Roth's theorem, we were able to shift and
rescale our progression to become an interval, and then iterate the argument until the density
increased above $1$. Unfortunately,
now we can't simply scale the progression $P$ to transform it into a {\it bona fide} interval, since then
square differences between elements in $A$ may no longer be square and vice versa. For
this reason, we will need another lemma that is suited for iteration.

\proclaim Lemma~\advthm.
Suppose that there is $N\in \NN$ and $A\subseteq [N]$ of density $\delta = |A|/N$ such that
$A-A$ does not contain any positive square numbers. Then there is a set $A'\subseteq [N^{1/4}]$
with $A' \ge (\delta + c'\delta^3) N^{1/4}$, where $c$ is an absolute constant, such that $A'-A'$
also does not contain any squares.

\proof By the previous lemma, there is a progression $P\subseteq [N]$ of length $\lfloor N^{1/3}$
and step size $1\le d \le 2N^{1/100}$ such that $|A\cap P|/|P| \ge \delta + c\delta^3$ for an absolute
constant $c$.
